Sampling in Decisions from Experience

### 2021-10-09 slides_modelers-meeting.Rmd
Add results and summary

---
title: "Sampling in Decisions from Experience"
subtitle: https://github.com/linushof/sampling-in-dfe
author: "Linus, Veronika & Thorsten"
date: "October 11, 2021"
output:
  beamer_presentation: default
bibliography: references_sampling-in-dfe.bib
csl: apa.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

pacman::p_load(tidyverse, 
               scico)
```

@@

- Does choice behavior reflect different sampling strategies?

@@

@@

```{r}
cols <- list(.default = col_double(),
             boundary = col_factor(),
             a = col_factor(),
             parameter = col_factor())
cpt <- read_csv("data/estimates/estimates_cpt.csv", col_types = cols)

@@

## plot curves
cpt_w %>% 
  filter(a != 1) %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = ep, y = w, color = s)) +
  geom_path(size = 1.2) + 
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) + 
  labs(title = "Probability Weighting Function",
       x ="Experienced Probability",
       y = "Subjective Experienced Probability", 
       color="Switching Probability") + 
  theme_minimal()
```

@@

```{r message=FALSE, warning=FALSE}
cols <- list(.default = col_double(),
             boundary = col_factor(),
             gamble = col_factor(),
             rare = col_factor(),
             agent = col_factor(),
             choice = col_factor())
choices <- read_csv("data/choices/choices.csv", col_types = cols)

@@

fr_rates %>% 
  ggplot(aes(s, rate, color = s)) + 
  facet_grid(type ~ fct_relevel(rare, "attractive", "none", "unattractive"), switch = "y") +
  geom_jitter(size = 3) + 
  scale_x_continuous(breaks = seq(0, 1, .1)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) +
  scale_color_scico(palette = "buda") + 
  labs(title = "False Response Rates", 
       x = "s", 
       y= "% False Responses", 
       color = "a") + 
  theme_minimal()
```

@@

```{r}
cpt_v <- cpt %>% 
  select(-c(gamma, delta, rho)) %>% 
  expand_grid(x = seq(0, 20, 2)) %>%  
  mutate(v = round(x^alpha, 2))

## plot curves
cpt_v %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = x, y = v)) +
  geom_path(aes(color = s)) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 20, 5)) +
  scale_y_continuous(breaks = seq(0, 20, 5)) + 
  labs(title = "Value Function",
       x ="Objective Value",
       y = "Subjective Value", 
       color="Switching Probability") + 
  theme_minimal()
```

>>>>

---
title: "Sampling in Decisions from Experience"
subtitle: https://github.com/linushof/sampling-in-dfe
author: "Linus, Veronika & Thorsten"
date: "October 11, 2021"
output:
  beamer_presentation: default
bibliography: references_sampling-in-dfe.bib
csl: apa.csl
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
pacman::p_load(tidyverse, 
               scico,
               ggpubr)
```

@@

- May samples be generated in such a way that they produce systematic choice patterns?

@@

## Method

- 60 choices between a 2-outcome prospect and a safe prospect
- 100 synthetic agents
- Parameterization of the sampling and decision process
  + Probability of switching between prospects
  + Number of comparisons
- Description of simulated DfE in Cumulative Prospect Theory

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
cols <- list(.default = col_double(),
             boundary = col_factor(),
             a = col_factor(),
             parameter = col_factor())
cpt <- read_csv("data/estimates/estimates_cpt.csv", col_types = cols)

@@

## plot curves
curve_w <- cpt_w %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = ep, y = w, color = s)) +
  geom_path() + 
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) + 
  labs(title = "Probability Weighting", 
       x ="Experienced Probability p",
       y = "w(p)", 
       color="Switching Probability") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="bottom")

gamma <- cpt %>% 
  group_by(boundary, a) %>% 
  ggplot(aes(x = s, y = gamma, color = s)) +
  geom_smooth(data = cpt %>% filter(a != 1), color = "grey") + 
  geom_jitter(size = 3, show.legend = FALSE) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) + 
  scale_y_continuous(breaks = seq(0, 2, .2)) +
  labs(title = expression(gamma),
       x ="",
       y = "") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))

delta <- cpt %>% 
  ggplot(aes(x = s, y = delta, color = s)) +
  geom_smooth(data = cpt %>% filter(a != 1), color = "grey") + 
  geom_jitter(size = 3, show.legend = FALSE) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) + 
  scale_y_continuous(limits = c(1, 2), breaks = seq(1, 2, .2)) +
  labs(title = expression(delta),
       x ="Switching Probability",
       y = "") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))

params <- ggarrange(gamma, delta, nrow = 2)
ggarrange(curve_w, params)
```

## Probability Weighting: Multiple Comparisons

```{r echo=FALSE}
cpt_w %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = ep, y = w, color = s)) +
  geom_path(size = 1.2) + 
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) + 
  labs(x ="Experienced Probability p",
       y = "w(p)", 
       color="Switching Probability") + 
  facet_wrap(~a, nrow = 2) +  
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="bottom")
```

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
cols <- list(.default = col_double(),
             boundary = col_factor(),
             gamble = col_factor(),
             rare = col_factor(),
             agent = col_factor(),
             choice = col_factor())
choices <- read_csv("data/choices/choices.csv", col_types = cols)

@@

fr_rates %>% 
  ggplot(aes(s, a, fill = rate)) + 
  geom_tile(colour="white", size=0.1) +
  scale_fill_scico(palette = "buda") + 
  facet_grid(type ~ fct_relevel(rare, "attractive", "none", "unattractive"), switch = "y") +
  scale_x_continuous(expand=c(0,0), breaks = seq(0, 1, .1)) +
  scale_y_continuous(expand=c(0,0), breaks = seq(1, 5, 1)) +
  labs(title = "Type of Rare Event", 
       x = "Switching Probability", 
       y= "Number of Comparisons", 
       fill = "% False Responses") + 
  theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
```

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
cpt_v <- cpt %>% 
  select(-c(gamma, delta, rho)) %>% 
  expand_grid(x = seq(0, 20, 2)) %>%  
  mutate(v = round(x^alpha, 2))

## plot curves
curve_v <- cpt_v %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = x, y = v)) +
  geom_path(aes(color = s)) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 20, 2)) +
  scale_y_continuous(breaks = seq(0, 20, 2)) + 
  labs(title = "Value Function",
       x ="Objective Value x",
       y = "v(x)", 
       color="Switching Probability") + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="bottom")

alpha <- cpt %>% 
  group_by(boundary, a) %>% 
  ggplot(aes(x = s, y = alpha, color = s)) +
  geom_smooth(data = cpt %>% filter(a != 1), color = "grey") + 
  geom_jitter(size = 3, show.legend = FALSE) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) + 
  scale_y_continuous(breaks = seq(0, 2, .1)) +
  labs(title = expression(alpha),
       x ="Switching Probability",
       y = "") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(curve_v, alpha)
```

## Value Function: Multiple Comparisons

```{r echo=FALSE, message=FALSE, warning=FALSE}
cpt_v %>% 
  group_by(boundary, a, s) %>% 
  ggplot(aes(x = x, y = v, color = s)) +
  geom_path(size = 1.2) + 
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 20, 2)) +
  scale_y_continuous(breaks = seq(0, 20, 2)) + 
  labs(x ="Objective Value x",
       y = "v(x)", 
       color="Switching Probability") + 
  facet_wrap(~a, nrow = 2) +  
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="bottom")
```

## Summary and Q&A

Random processes underlying DfE ...

- ... can be modeled according to probability theory

- ... can be integrated into current decision theories

- ... could be shaped by sampling and decision strategies

- ... could explain empirically observed choice patterns

I happily take your questions, comments and critique.
